Tulai

### 2023-09-19 paper/paper_tulai.qmd
move chunks to make summary_table_area before we use it
exploring PCA

pca_data <- sum_depth_sheet %>% 
  select(lithic_dens, core_freq, blank_freq, retouch_freq, tool_diversity) %>%
  na.omit()  # Remove rows with NA values

@@

pca_data <- sum_depth_sheet[, c('lithic_dens', 'core_freq', 'blank_freq', 'retouch_freq', 'tool_diversity')]

@@

```{r}
#Plot: Contribution of variables for each of the four relevant PCA dimensions
library(factoextra)
library(ggplot2)

#save contribution plots into one objects
p1 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 1, 
                   top = 10)
p2 <- fviz_contrib(res.pca, 
                   choice = "var",
                   axes = 2, 
                   top = 10)
p3 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 3, 
                   top = 10)
p4 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 4, 
                   top = 10)
p5 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 5, 
                   top = 10)

#combine individual plots into one plot 
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, ncol = 2)

@@

print("Significant contributions to Dim5:")
print(sorted_high_contrib_dim5)

```{r}
# screeplot; inspect distribution of PCs
library(factoextra)  
fviz_screeplot(res.pca)
```

```{r}
# Inspect eigenvalues
print(res.pca$eig)

@@

# Display the table
htmltools::browsable(kable_output)

```{r}
# Visualize biplot:1
fviz_pca_biplot(res.pca,
                axes = c(1, 2),
                labelsize = 2
                )

rownames(res.pca$ind$coord) <- sum_depth_sheet$label

@@

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2
                              )

# Modify row names of coordinates
rownames(res.pca$ind$coord) <- sum_depth_sheet$label

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = as.data.frame(res.pca$ind$coord),
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = rownames(res.pca$ind$coord)
                                       ),
                                   box.padding = 0.5,
                                   point.padding = 0.5,
                                   size = 2
                                  )

@@

```{r}
# Visualize biplot: 2
fviz_pca_biplot(res.pca,
                axes = c(3, 4),
                labelsize = 2
                )
rownames(res.pca$ind$coord) <- sum_depth_sheet$label

>>>>

@@

sum_depth_sheet_subset <- 
 sum_depth_sheet %>%
  drop_na(tool_diversity) #%>% 
  #filter(!label %in% c("tp1_c_3_a",
 #                      "d1_a_2_b",
  #                     "d1_c_3_b"))

pca_data <- 
sum_depth_sheet_subset %>% 
  select(lithic_dens, 
         core_freq, 
         blank_freq, 
         retouch_freq, 
         tool_diversity)

@@

```{r}
#Plot: Contribution of variables for each of the four relevant PCA dimensions

#save contribution plots into one objects

contribution_plots <- 
map(1:5, ~fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = .x, 
                   top = 10))

#combine individual plots into one plot 
library(cowplot)
plot_grid(plotlist = contribution_plots,
                   ncol = 2)

@@

print("Significant contributions to Dim5:")
print(sorted_high_contrib_dim5)
```

```{r}
# screeplot; inspect distribution of PCs
fviz_screeplot(res.pca)
```

```{r}
# Inspect eigenvalues
res.pca$eig
```

@@

# Display the table
kable_output

@@

rownames(res.pca$ind$coord) <- sum_depth_sheet_subset$label

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2,
                               label = "var",
                               addEllipses = TRUE,
                               col.ind = sum_depth_sheet_subset$area
                              )

# Modify row names of coordinates
rownames(res.pca$ind$coord) <- sum_depth_sheet_subset$label

pca_label_text <- 
res.pca$ind$coord %>% 
  as_tibble() %>% 
  mutate(labels = sum_depth_sheet_subset$label,
         area =   sum_depth_sheet_subset$area)

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = pca_label_text,
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = labels,
                                       colour = area),
                                   segment.color = "grey80",
                                   size = 2,
                                   force = 10
                                  )

@@

```{r}
# Visualize biplot: 2
fviz_pca_biplot(res.pca,
                axes = c(3, 4),
                labelsize = 2
                )
rownames(res.pca$ind$coord) <- sum_depth_sheet$label
```
